---
title: Practical Machine Learning Project
output:
  html_document:
    fig_height: 9
    fig_width: 9
---

## Introduction  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

  

## Adding libraries and downloading files
```{r, cache = T}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
library(e1071)

trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./pml-training.csv"
testFile  <- "./pml-testing.csv"

if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile)
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile)
}

```

### Read the Data
  
```{r, cache = T}
trainData <- read.csv("./pml-training.csv")
testData <- read.csv("./pml-testing.csv")
dim(trainData)
dim(testData)
```

### Work with data
We will clean data - remove NA values, delete unnecesarry columns

```{r, cache = T}
trainData <- trainData[, colSums(is.na(trainData)) == 0] 
testData <- testData[, colSums(is.na(testData)) == 0]
classe <- trainData$classe
trainRemove <- grepl("^X|timestamp|window", names(trainData))
trainData <- trainData[, !trainRemove]
trainCleaned <- trainData[, sapply(trainData, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testData))
testData <- testData[, !testRemove]
testCleaned <- testData[, sapply(testData, is.numeric)]
```  


### Data split
Split data into training - 70% and validation - 30% sets.  
```{r, cache = T}
set.seed(22222)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling
We will use Random Forest, as it selects important variables. 
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
We then check result on our test  
```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```
```{r, cache = T}
accuracy <- postResample(predictRf, testData$classe)
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```
Estimated accuracy is higher than 99%

## Prediction on test set
We apply trained model to downloaded test data 
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

## Correlation Matrix Visualization  
```{r, cache = T}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```
## Decision Tree Visualization
```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) 